FAIR Theory: Applying Open Science Principles to the Construction and Iterative Improvement of Scientific Theories

### 2024-12-13 manuscript.Rmd
Update manuscript.Rmd
add a sentence to tell people: don't panic

Concerns about the state of theory in the psychological literature revolve around two issues: theory formalization and theory (re-)use.
Greater formalization increases theories' *empirical content* [REF] as it forces researchers to use precise statements, for example by specifying exact functional forms for relations or forces them to specify processes  that would otherwise get away with their vague formulation in verbal theories.
For example, the phonological loop in Baddeley's verbal description of his working memory model allows at least 144 different implementations, for example, depending on how decay rate, recall success, or rehearsal sequence are precisely implemented [@lewandowsky2010computational].
More precise theories makes them easier to falsify,
which necessitates revising them,
thus advancing our principled understanding of the phenomena they describe.
FAIR Theory does not require theories to be formal,
and formal theory can be represented in a way that is not FAIR.
For example, it is possible to represent a collection of verbal propositions (perhaps derived through qualitative research) as a FAIR theory.
Conversely, a directed acyclic graph (DAG) is a type of formal theory,
but if it is embedded within a journal article as a bitmap image without any key words to help search engines index that article as a theory paper,
then this formal theory is not FAIR.
FAIR Theory is thus consistent with, but does not require, formal theory (also see *Accessibility*).

>>>>

Concerns about the state of theory in the psychological literature revolve around two issues: theory formalization and theory (re-)use.
Greater formalization increases theories' *empirical content* [REF] as it forces researchers to use precise statements, for example by specifying exact functional forms for relations or forces them to specify processes  that would otherwise get away with their vague formulation in verbal theories.
For example, the phonological loop in Baddeley's verbal description of his working memory model allows at least 144 different implementations, for example, depending on how decay rate, recall success, or rehearsal sequence are precisely implemented [@lewandowsky2010computational].
More precise theories makes them easier to falsify,
which necessitates revising them,
thus advancing our principled understanding of the phenomena they describe.
FAIR Theory does not require theories to be formal,
and formal theory can be represented in a way that is not FAIR.
It is therefore important for us to emphasize that FAIR Theory imposes no restrictions on researchers regarding the manner in which theories are derived and written down. 
The guidelines introduced by FAIR Theory primarily pertain to how theories are documented and shared in digital environments, with the aim of enhancing their reusability and extensibility.
For example, it is possible to represent a collection of verbal propositions (perhaps derived through qualitative research) as a FAIR theory.
Conversely, a directed acyclic graph (DAG) is a type of formal theory,
but if it is embedded within a journal article as a bitmap image without any key words to help search engines index that article as a theory paper,
then this formal theory is not FAIR.
FAIR Theory is thus consistent with, but does not require, formal theory (also see *Accessibility*).
